The Effects of the 2023 Rule Changes in Major League Baseball

### 2023-11-10 manuscript/paper.tex
finished methods section of manuscript

Use this section to present the methodologies that will generate results by
analyzing the data.

>>>>

The introduction of these rule changes in the MLB provides a new and exciting
opportunity to study how they impacted various aspects of the game. This study
mainly focuses on comparing baseball statistics from the 2023 season versus the
2022 season to limit potential confounding factors. If we instead compared the
post-rule change season to say the last 10 MLB seasons, observed differences
may not be solely due to the 2023 rule changes, but rather from the effects of
drastic changes in the game between now and then. For instance, there were other
rule changes made in the last 10 years, such as the National League (NL) adopting
the designated hitter in 2022, relief pitchers having to pitch to at least 3 batters,
etc., that impacted statistics around the league. The 2020 season was also
drastically different due to COVID-19, with less games being played and no fans in
the stands. The ``juiced balls'' conspiracy was also running rampant in the last
decade, with evidence to show that in some seasons, players were playing with
baseballs that traveled further and harder off the bat. Offensive and defensive
philosophy has also seen historical changes, with defensive shifting becoming
more and more prevalent and teams diving more and more into the ``home run or
bust''/launch angle theory with batting. However, 2022 and 2023 were fairly similar
in nature all things considered with the obvious change being the new rules implemented,
which we will test to see how they affected statistics in 2023.

Comparisons between the 2023 and 2022 seasons were made on offensive metrics such
as batting average, stolen bases, and slugging percentage, along with game length.
Comparing the difference in these metrics between seasons will reveal valuable insights
into understanding how the rule changes impacted the game.

The ten following specific research questions will test the effects of the 2023
MLB rule changes on the game of baseball:

\begin{enumerate}
  \def\labelenumi{\arabic{enumi}.}
  \item
    Is the average game length shorter in 2023 compared to the previous 25 years?
  \item
    Are players stealing more bases in 2023 than they did in 2022?
  \item
    Is the strikeout rate different in 2023 compared to 2022?
  \item
    Is the walk rate different in 2023 compared to 2022?
  \item
    Is the batting average different in 2023 compared to 2022?
  \item
    Is the slugging percentage different in 2023 compared to 2022?
  \item
    Is the batting average on balls in play different in 2023 compared to 2022?
  \item 
    Is the expected batting average different in 2023 compared to 2022?
  \item
    Is the number of hits on ground balls greater in 2023 compared to 2022?
  \item
    Is the number of ground balls into double plays different in 2023 compared to 2022?
\end{enumerate}

All ten questions can be tested using a standard two-sample comparison with the \(t\)-test.
The \(z\)-test was considered as done in \citet{Price}, but we do not know the population standard deviation
in this case. The \(t\)-test assumes the data is normally distributed so to account
for this, we also conducted the nonparametric, distribution free, Mann Whitney U
Test (also known as the Wilcoxon Rank Sum Test) to be sure of our results. Also,
notice how some of these research questions call for a one-sided hypothesis test,
like the number of hits on ground balls, because we intuitively expect this number
to be greater in 2023 than in 2022 due to the limit of the defensive shift. Other
metrics that we were not sure about whether they would increase or decrease were
tested using a two-sided hypothesis test, simply testing for a significance in the
overall difference from season to season. Other methods such as regression were
considered for this study as well, but we are primarily interested in comparing
means and assessing the effect of a single variable, so a \(t\)-test is more appropriate
and straightforward.
